Geospatial and Temporal Analysis of Toronto's Cycling Infrastructure

### 2024-09-24 paper/paper.qmd
paper first draft-minor update

@fig-comfort-one provides a breakdown of Toronto’s bikeway network by comfort level and upgrade status. @fig-comfort-one-1 shows the distribution of all bikeways based on comfort level, with the majority of bikeways classified as either Low Comfort (41.6%) or Moderate Comfort (41.1%), while High Comfort lanes constitute a smaller portion (17.3%). This indicates that while Toronto has invested in a significant number of bike lanes, most of them fall within lower comfort categories, suggesting that further improvements, particularly in upgrading existing lanes, are needed to promote safer and more comfortable cycling experiences.

@@

@fig-comfort-one-2, @fig-comfort-two-1, and @fig-comfort-two-2 offer a deeper look into the upgrade status for each comfort level. For Low Comfort lanes, only 22.69% have been upgraded, while 77.31% remain in their original state. Similarly, for Moderate Comfort bikeways, 23.38% have been upgraded, leaving 76.62% untouched. Interestingly, High Comfort lanes have the lowest proportion of upgrades, with only 18.76% improved, despite representing the most comfortable and potentially most desirable biking infrastructure. These trends highlight a potential gap in Toronto's focus on maintaining and upgrading its most comfortable and accessible cycling lanes, suggesting a need for more targeted investments in high-comfort bikeways to encourage broader use of cycling for transportation.

@@

Comfort level classification is crucial for determining which segments of the population feel confident using the cycling network. Low Comfort lanes, including painted bike lanes and shared roadways, provide minimal physical separation from vehicular traffic. This limited protection often discourages risk-averse cyclists, particularly in areas with high traffic volumes. While such lanes may be acceptable in quieter residential areas, they are often avoided in busy urban environments. Unfortunately, as @fig-comfort-one-2 shows, 77.31% of these Low Comfort lanes remain unupgraded, revealing a substantial gap in the network's ability to offer safe and comfortable conditions for all cyclists.

>>>>

@fig-comfort-one provides a breakdown of Toronto’s bikeway network by comfort level and upgrade status. @fig-comfort-one-1 shows the distribution of all bikeways based on comfort level, with the majority of bikeways classified as either Low Comfort (40.88%) or Moderate Comfort (41.75%), while High Comfort lanes constitute a smaller portion (17.36%). This indicates that while Toronto has invested in a significant number of bike lanes, most of them fall within lower comfort categories, suggesting that further improvements, particularly in upgrading existing lanes, are needed to promote safer and more comfortable cycling experiences.

@@

@fig-comfort-one-2, @fig-comfort-two-1, and @fig-comfort-two-2 offer a deeper look into the upgrade status for each comfort level. For Low Comfort lanes, only 22.26% have been upgraded, while 77.74% remain in their original state. Similarly, for Moderate Comfort bikeways, 23.38% have been upgraded, leaving 76.62% untouched. Interestingly, High Comfort lanes have the lowest proportion of upgrades, with only 18.76% improved, despite representing the most comfortable and potentially most desirable biking infrastructure. These trends highlight a potential gap in Toronto's focus on maintaining and upgrading its most comfortable and accessible cycling lanes, suggesting a need for more targeted investments in high-comfort bikeways to encourage broader use of cycling for transportation.

@@

Comfort level classification is crucial for determining which segments of the population feel confident using the cycling network. Low Comfort lanes, including painted bike lanes and shared roadways, provide minimal physical separation from vehicular traffic. This limited protection often discourages risk-averse cyclists, particularly in areas with high traffic volumes. While such lanes may be acceptable in quieter residential areas, they are often avoided in busy urban environments. Unfortunately, as @fig-comfort-one-2 shows, 77.74% of these Low Comfort lanes remain unupgraded, revealing a substantial gap in the network's ability to offer safe and comfortable conditions for all cyclists.
